A new breeding crossover approach for evolutionary algorithms

### 2023-09-14 paper.tex
First Draft

    % Please add the following required packages to your document preamble:
    % \usepackage{booktabs}
    % \usepackage[table,xcdraw]{xcolor}
    % If you use beamer only pass "xcolor=table" option, i.e. \documentclass[xcolor=table]{beamer}
    \begin{table}[]
    \scriptsize
    \centering
    \caption{Mathematical benchmark functions introduced in the Competition on Evolutionary Computation for the 2017 edition (CEC-2017).}\label{tab.benchmark_functions}    
    \begin{tabular}{@{}cllcl@{}}
    \toprule
    \multicolumn{5}{c}{\textbf{CEC-2017 Mathematical Benchmark Functions}} \\ \midrule
    \textbf{Fx} & \textbf{Function Name} &  & \textbf{Fx} & \textbf{Function Name} \\
    F1 & Shifted and Rotated   Bent Cigar Function &  & F16 & Hybrid function 6 \\
    F2 & Shifted and Rotated   Sum of Different Power Function &  & F17 & Hybrid function 7 \\
    F3 & Shifted and Rotated   Zakharov Function &  & F18 & Hybrid function 8 \\
    F4 & Shifted and Rotated   Rosenbrock's Function &  & F19 & Hybrid function 9 \\
    F5 & Shifted and Rotated   Rastrigin's Function &  & F20 & Hybrid function 10 \\
    F6 & Shifted and Rotated   Schaffer F7 Function &  & F21 & Composition function   1 \\
    F7 & Shifted and Rotated   Lunacek Bi-Rastrigin's Function &  & F22 & Composition function   2 \\
    F8 & Shifted and Rotated   Non-Continuous Rastrigin's Function &  & F23 & Composition function   3 \\
    F9 & Shifted and Rotated   Levy Function &  & F24 & Composition function   4 \\
    F10 & Shifted and Rotated   Schwefel's Function &  & F25 & Composition function   5 \\
    F11 & Hybrid function 1 &  & F26 & Composition function   6 \\
    F12 & Hybrid function 2 &  & F27 & Composition function   7 \\
    F13 & Hybrid function 3 &  & F28 & Composition function   8 \\
    F14 & Hybrid function 4 &  & F29 & Composition function   9 \\
    F15 & Hybrid function 5 &  & F30 & Composition function   10 \\ \bottomrule
    \end{tabular}
    \end{table}

@@

    % Please add the following required packages to your document preamble:
    % \usepackage{booktabs}
    % \usepackage[table,xcdraw]{xcolor}
    % If you use beamer only pass "xcolor=table" option, i.e. \documentclass[xcolor=table]{beamer}
    \begin{table}[]
    \scriptsize
    \centering
    \caption{Animal Life Cycle Algorithm (ALCA) General Configuration.}\label{tab.general_configuration}    
    \begin{tabular}{@{}ll@{}}
    \toprule
    \multicolumn{2}{l}{\textbf{ALCA   General Configuration}} \\ \midrule
    \textbf{Function name} & \textbf{ALL} \\
    Population & 500 \\
    Evaluations & 10,000 * Dim \\
    Crossover rate & 100 \\
    Mutation rate & 7 \\
    Max age & 5 \\
    Tournament rep. & 100 \\
    Sample size & 20 \\
    Base approval & 80 \\
    Goal approval & 200 \\ \bottomrule
    \end{tabular}
    \end{table}

@@

    For each experiment, we ran fifty-one independent executions per CEC-2017 mathematical benchmark function and dimensions specified. We recorded the following results: best-found error average, standard deviation, and calculated the statistical Z-Test value. The labels used on the results of our summarized results tables are the following:

@@

    To find and determine the best alternative finalized with a statistical Z-Test with a 95 percent confidence level. Here are the results statistics values: Table 3 shows the condensed ALCA evaluation results for the CEC-2017 functions running on ten dimensions (D) for alternatives One Point and Continuous Range crossover; Table 4 displays the condensed ALCA evaluation results for the CEC-2017 mathematical operations running on thirty D for the same crossover alternatives.

    We performed the statistical Z-Test for both crossover alternatives with a 95 percent confidence level. When the One Point passes the test, it is highlighted in green color. When the Continuous Range passes the statistical test, it is highlighted in blue. Where we did not find enough evidence to declare one choice as the best alternative, no Z value was highlighted.

@@

    STOP HERE --- A new beginning (this is an upload test).

    Imagine being an astronaut on a mission repairing a space station in the
    middle of space. Suddenly, a block of asteroids hits the craft, and you
    lose your oxygen tank while the explosion pushes you away. From afar, you
    see the area where your tank might be, covered with something similar to a
    gas leaked from the station that blocks all visibility. Finding the tank is
    the only hope to return to your main ship some miles away. The oxygen
    reserve from your suit is running out, and there are only a couple of
    minutes left. Now hold your breath in real life and think, what strategy
    will you choose to search for your tank? We repeated this mental exercise
    several times until we found the best solution (or blacked out).

    If your life depended on it, would you choose a Genetic Algorithm, a
    variant of a Particle Swarm Optimization Algorithm, or another? We searched
    for a way to mix both, at least in some way, take the main idea or concepts
    from them and make them work together. We strongly believe it is possible
    to combine the features, as previous work has proven to obtain outstanding
    results \cite{garcia2015evospace,garcia2021event,valdez2021container,valdez2021swarm}.
    Our goal was to create a swarm out of the best individuals found in
    evolution. We probably have a long way to go, but at least we are getting
    promising results. From our experiments, all six Classic Benchmark
    Functions for Optimization (Ackley, Bohachevsky, Griewank, Rastrigin,
    Sphere, and Rosenbrock) proved the same behavior for the Animal Life Cycle
    Algorithm: finding the expected solution in a reduced number of evaluations
    and less execution time.

@@

    The Animal Life Cycle Algorithm allows for multiple parameters'
    fine-tuning, facilitating the freedom to experiment with different
    alternatives to restart the population and solve extinction caused by
    nature's pressure, which will impact how quickly, and the quality of its
    found solution. We compared the results obtained with the use of historical
    elite projection versus other alternatives, using classic benchmark
    functions for optimization for comparison, where it showed favorable and
    promising results.

    As the nature of the challenge increases, it is indispensable to have an
    elastic, scalable, and fault-tolerant model designed with cloud
    collaboration processes, and asynchronous communication. We have proven
    that it is possible to evolve a population, using a distributed, parallel,
    and asynchronous strategy, inspired by the Genetic Algorithm (GA). To
    further validate this work, we could use some more demanding (or control)
    problem that requires calculating real numbers
    \cite{stanley2002evolving,miikkulainen2019evolving}. We implemented the
    algorithm using Docker containers.

>>>>

    \begin{table}[]
    \scriptsize
    \centering
    \caption{Mathematical benchmark functions introduced in the Competition on Evolutionary Computation for the 2017 edition (CEC-2017).}\label{tab.benchmark_functions}    
    \begin{tabular}{@{}cllcl@{}}
    \toprule
    \multicolumn{5}{c}{\textbf{CEC-2017 Mathematical Benchmark Functions}} \\ \midrule
    \textbf{Fx} & \textbf{Function Name} &  & \textbf{Fx} & \textbf{Function Name} \\
    F1 & Shifted and Rotated   Bent Cigar Function &  & F16 & Hybrid function 6 \\
    F2 & Shifted and Rotated   Sum of Different Power Function &  & F17 & Hybrid function 7 \\
    F3 & Shifted and Rotated   Zakharov Function &  & F18 & Hybrid function 8 \\
    F4 & Shifted and Rotated   Rosenbrock's Function &  & F19 & Hybrid function 9 \\
    F5 & Shifted and Rotated   Rastrigin's Function &  & F20 & Hybrid function 10 \\
    F6 & Shifted and Rotated   Schaffer F7 Function &  & F21 & Composition function   1 \\
    F7 & Shifted and Rotated   Lunacek Bi-Rastrigin's Function &  & F22 & Composition function   2 \\
    F8 & Shifted and Rotated   Non-Continuous Rastrigin's Function &  & F23 & Composition function   3 \\
    F9 & Shifted and Rotated   Levy Function &  & F24 & Composition function   4 \\
    F10 & Shifted and Rotated   Schwefel's Function &  & F25 & Composition function   5 \\
    F11 & Hybrid function 1 &  & F26 & Composition function   6 \\
    F12 & Hybrid function 2 &  & F27 & Composition function   7 \\
    F13 & Hybrid function 3 &  & F28 & Composition function   8 \\
    F14 & Hybrid function 4 &  & F29 & Composition function   9 \\
    F15 & Hybrid function 5 &  & F30 & Composition function   10 \\ \bottomrule
    \end{tabular}
    \end{table}

@@

    \begin{table}[]
    \scriptsize
    \centering
    \caption{Animal Life Cycle Algorithm (ALCA) General Configuration.}\label{tab.general_configuration}    
    \begin{tabular}{@{}ll@{}}
    \toprule
    \multicolumn{2}{l}{\textbf{ALCA   General Configuration}} \\ \midrule
    \textbf{Function name} & \textbf{ALL} \\
    Population & 500 \\
    Evaluations & 10,000 * Dim \\
    Crossover rate & 100 \\
    Mutation rate & 7 \\
    Max age & 5 \\
    Tournament rep. & 100 \\
    Sample size & 20 \\
    Base approval & 80 \\
    Goal approval & 200 \\ \bottomrule
    \end{tabular}
    \end{table}

@@

    For each experiment, we ran fifty-one independent executions per CEC-2017 mathematical benchmark function and dimensions specified. We recorded the following results: best-found error average and standard deviation to calculate the statistical Z-Test value. The labels used on the results of our summarized results tables are the following:

@@

    To find and determine the best alternative finalized with a statistical Z-Test with a ninety-five percent confidence level. Here are the results statistics values: Table~\ref{tab.condensed_evaluation_10D} shows the condensed ALCA evaluation results for the CEC-2017 functions running on ten dimensions (D) for alternatives One Point and Continuous Range crossover; Table~\ref{tab.condensed_evaluation_30D} displays the condensed ALCA evaluation results for the CEC-2017 mathematical operations running on thirty D for the same crossover alternatives.

    \begin{table}[]
    \scriptsize
    \centering
    \caption{Condensed ALCA evaluation results for the CEC-2017 functions running on ten dimensions for alternatives One Point and Continuous Range crossover.}\label{tab.condensed_evaluation_10D}
    \begin{tabular}{@{}lllllllll@{}}
    \toprule
    \multicolumn{9}{c}{\textbf{ALCA . CEC-2017 . 10D}} \\ \midrule
    Crossover &  & \multicolumn{3}{c}{One Point Xover} &  & \multicolumn{3}{c}{Cont. Range Xover} \\
    Fx &  & Mean & Std. Dev. & Z &  & Mean & Std. Dev. & Z \\
    F1 &  & 2.14E+03 & 2.14E+03 & 5.42E+00 &  & 4.43E+02 & 6.46E+02 & \textbf{-5.42E+00} \\
    F2 &  & 1.81E-03 & 2.63E-03 & 2.38E+00 &  & 8.54E-04 & 1.16E-03 & \textbf{-2.38E+00} \\
    F3 &  & 4.19E-04 & 1.16E-03 & 2.49E+00 &  & 1.33E-05 & 2.20E-05 & \textbf{-2.49E+00} \\
    F4 &  & 6.00E+00 & 1.31E+01 & 2.00E+00 &  & 2.31E+00 & 1.78E+00 & \textbf{-2.00E+00} \\
    F5 &  & 1.53E+01 & 7.06E+00 & 2.45E+00 &  & 1.24E+01 & 4.14E+00 & \textbf{-2.45E+00} \\
    F6 &  & 6.05E-04 & 9.22E-04 & -1.43E+00 &  & 2.53E-03 & 9.61E-03 & 1.43E+00 \\
    F7 &  & 3.15E+01 & 9.06E+00 & 2.26E+00 &  & 2.76E+01 & 8.38E+00 & \textbf{-2.26E+00} \\
    F8 &  & 1.78E+01 & 7.38E+00 & 8.00E+00 &  & 8.62E+00 & 3.53E+00 & \textbf{-8.00E+00} \\
    F9 &  & 2.02E+01 & 2.24E+01 & 1.85E+00 &  & 1.25E+01 & 1.96E+01 & \textbf{-1.85E+00} \\
    F10 &  & 4.96E+02 & 2.26E+02 & \textbf{-2.02E+00} &  & 5.94E+02 & 2.63E+02 & 2.02E+00 \\
    F11 &  & 1.16E+01 & 6.01E+00 & \textbf{-2.44E+00} &  & 1.49E+01 & 7.48E+00 & 2.44E+00 \\
    F12 &  & 1.70E+04 & 1.99E+04 & 2.00E+00 &  & 1.10E+04 & 8.21E+03 & \textbf{-2.00E+00} \\
    F13 &  & 7.39E+03 & 8.16E+03 & 1.08E+00 &  & 5.94E+03 & 5.03E+03 & -1.08E+00 \\
    F14 &  & 3.24E+02 & 6.05E+02 & 4.65E-01 &  & 2.67E+02 & 6.34E+02 & -4.65E-01 \\
    F15 &  & 5.70E+02 & 1.31E+03 & 1.25E+00 &  & 3.07E+02 & 7.46E+02 & -1.25E+00 \\
    F16 &  & 1.56E+02 & 1.28E+02 & 5.05E-01 &  & 1.43E+02 & 1.18E+02 & -5.05E-01 \\
    F17 &  & 3.17E+01 & 3.65E+01 & 1.81E+00 &  & 2.11E+01 & 1.98E+01 & \textbf{-1.81E+00} \\
    F18 &  & 2.26E+03 & 2.88E+03 & 9.03E-01 &  & 1.73E+03 & 3.06E+03 & -9.03E-01 \\
    F19 &  & 1.88E+03 & 3.24E+03 & 4.11E-01 &  & 1.63E+03 & 2.95E+03 & -4.11E-01 \\
    F20 &  & 9.50E+00 & 8.32E+00 & -1.35E+00 &  & 1.20E+01 & 9.92E+00 & 1.35E+00 \\
    F21 &  & 1.15E+02 & 3.86E+01 & 4.93E-02 &  & 1.15E+02 & 3.58E+01 & -4.93E-02 \\
    F22 &  & 1.00E+02 & 2.07E+01 & \textbf{-2.57E+00} &  & 1.08E+02 & 8.70E+00 & 2.57E+00 \\
    F23 &  & 3.21E+02 & 7.31E+00 & \textbf{-2.53E+00} &  & 3.25E+02 & 8.91E+00 & 2.53E+00 \\
    F24 &  & 3.28E+02 & 8.06E+01 & 5.36E+00 &  & 2.15E+02 & 1.28E+02 & \textbf{-5.36E+00} \\
    F25 &  & 4.35E+02 & 2.61E+01 & 1.39E+00 &  & 4.28E+02 & 2.41E+01 & -1.39E+00 \\
    F26 &  & 3.99E+02 & 1.25E+02 & 7.55E-01 &  & 3.82E+02 & 9.35E+01 & -7.55E-01 \\
    F27 &  & 3.88E+02 & 1.03E+01 & -5.43E-01 &  & 3.90E+02 & 1.13E+01 & 5.43E-01 \\
    F28 &  & 4.11E+02 & 7.92E+01 & 1.93E+00 &  & 3.80E+02 & 8.46E+01 & \textbf{-1.93E+00} \\
    F29 &  & 3.00E+02 & 4.32E+01 & 7.44E-01 &  & 2.94E+02 & 3.30E+01 & -7.44E-01 \\
    F30 &  & 3.21E+02 & 1.52E+02 & -7.14E-01 &  & 3.44E+02 & 1.81E+02 & 7.14E-01 \\
    \textbf{Total (W)} &  & \multicolumn{2}{c}{\textbf{One Point Xover}} & \multicolumn{1}{c}{\textbf{4 of 30}} &  & \multicolumn{2}{c}{\textbf{Cont. Range Xover}} & \multicolumn{1}{c}{\textbf{12 of 30}} \\ \bottomrule
    \end{tabular}
    \end{table}

@@

    We performed the statistical Z-Test for both crossover alternatives with a 95 percent confidence level. When the One Point passes the test, the Z value is highlighted in bold on the left column. When the Continuous Range passes the statistical test, the Z value is highlighted in bold on the right column. Where we did not find enough evidence to declare one choice as the best alternative, no value was highlighted.

    \begin{table}[]
    \scriptsize
    \centering
    \caption{Condensed ALCA evaluation results for the CEC-2017 functions running on thirty dimensions for alternatives One Point and Continuous Range crossover.}\label{tab.condensed_evaluation_30D}    
    \begin{tabular}{@{}lllllllll@{}}
    \toprule
    \multicolumn{9}{c}{\textbf{ALCA . CEC-2017 . 30D}} \\ \midrule
    Crossover &  & \multicolumn{3}{c}{One Point Xover} &  & \multicolumn{3}{c}{Cont. Range Xover} \\
    Fx &  & Mean & Std. Dev. & Z &  & Mean & Std. Dev. & Z \\
    F1 &  & 4.92E+03 & 6.21E+03 & 1.60E+00 &  & 3.35E+03 & 3.33E+03 & -1.60E+00 \\
    F2 &  & 9.21E+06 & 4.77E+07 & 1.37E+00 &  & 6.96E+04 & 2.50E+05 & -1.37E+00 \\
    F3 &  & 1.17E+02 & 1.18E+02 & 4.33E+00 &  & 4.16E+01 & 3.85E+01 & \textbf{-4.33E+00} \\
    F4 &  & 7.17E+01 & 3.14E+01 & -5.29E-01 &  & 7.46E+01 & 2.52E+01 & 5.29E-01 \\
    F5 &  & 9.38E+01 & 2.41E+01 & 4.61E-01 &  & 9.18E+01 & 2.04E+01 & -4.61E-01 \\
    F6 &  & 4.67E-03 & 1.02E-02 & \textbf{-2.42E+00} &  & 2.28E-02 & 5.26E-02 & 2.42E+00 \\
    F7 &  & 1.55E+02 & 3.04E+01 & -3.88E-01 &  & 1.58E+02 & 3.44E+01 & 3.88E-01 \\
    F8 &  & 9.94E+01 & 2.43E+01 & 6.82E+00 &  & 7.29E+01 & 1.32E+01 & \textbf{-6.82E+00} \\
    F9 &  & 1.37E+03 & 7.21E+02 & 4.60E+00 &  & 8.19E+02 & 4.64E+02 & \textbf{-4.60E+00} \\
    F10 &  & 2.79E+03 & 5.59E+02 & \textbf{-2.22E+00} &  & 3.02E+03 & 4.63E+02 & 2.22E+00 \\
    F11 &  & 7.31E+01 & 2.72E+01 & -1.09E+00 &  & 7.90E+01 & 2.75E+01 & 1.09E+00 \\
    F12 &  & 2.18E+05 & 1.94E+05 & 4.61E+00 &  & 8.73E+04 & 5.95E+04 & \textbf{-4.61E+00} \\
    F13 &  & 1.95E+04 & 2.73E+04 & 2.09E+00 &  & 1.10E+04 & 8.96E+03 & \textbf{-2.09E+00} \\
    F14 &  & 2.22E+04 & 2.14E+04 & 2.83E+00 &  & 1.27E+04 & 1.06E+04 & \textbf{-2.83E+00} \\
    F15 &  & 1.25E+04 & 1.41E+04 & 4.96E+00 &  & 2.37E+03 & 3.82E+03 & \textbf{-4.96E+00} \\
    F16 &  & 1.26E+03 & 3.45E+02 & 1.49E-01 &  & 1.26E+03 & 2.49E+02 & -1.49E-01 \\
    F17 &  & 6.02E+02 & 1.76E+02 & 1.17E+00 &  & 5.54E+02 & 2.36E+02 & -1.17E+00 \\
    F18 &  & 1.29E+05 & 1.17E+05 & 2.92E+00 &  & 7.44E+04 & 6.27E+04 & \textbf{-2.92E+00} \\
    F19 &  & 9.78E+03 & 1.16E+04 & 3.04E+00 &  & 4.38E+03 & 5.08E+03 & \textbf{-3.04E+00} \\
    F20 &  & 4.83E+02 & 1.69E+02 & 1.02E-01 &  & 4.79E+02 & 2.03E+02 & -1.02E-01 \\
    F21 &  & 3.03E+02 & 2.44E+01 & 4.40E+00 &  & 2.85E+02 & 1.62E+01 & \textbf{-4.40E+00} \\
    F22 &  & 1.76E+03 & 1.72E+03 & 6.07E+00 &  & 2.16E+02 & 5.84E+02 & \textbf{-6.07E+00} \\
    F23 &  & 4.62E+02 & 3.06E+01 & \textbf{-5.27E+00} &  & 5.05E+02 & 4.97E+01 & 5.27E+00 \\
    F24 &  & 6.45E+02 & 7.43E+01 & \textbf{-2.31E+00} &  & 6.78E+02 & 7.03E+01 & 2.31E+00 \\
    F25 &  & 3.85E+02 & 1.49E+01 & \textbf{-3.06E+00} &  & 3.97E+02 & 2.40E+01 & 3.06E+00 \\
    F26 &  & 2.40E+03 & 4.88E+02 & -1.23E+00 &  & 2.61E+03 & 1.17E+03 & 1.23E+00 \\
    F27 &  & 4.94E+02 & 1.23E+01 & 2.98E-01 &  & 4.93E+02 & 1.95E+01 & -2.98E-01 \\
    F28 &  & 4.63E+02 & 4.29E+01 & 9.26E+00 &  & 3.98E+02 & 2.58E+01 & \textbf{-9.26E+00} \\
    F29 &  & 9.22E+02 & 2.36E+02 & 1.55E+00 &  & 8.56E+02 & 2.00E+02 & -1.55E+00 \\
    F30 &  & 3.53E+03 & 4.15E+03 & 4.09E+00 &  & 1.08E+03 & 1.04E+03 & \textbf{-4.09E+00} \\
    \multicolumn{1}{c}{\textbf{Total (W)}} & \multicolumn{1}{c}{\textbf{}} & \multicolumn{2}{c}{\textbf{One Point Xover}} & \multicolumn{1}{c}{\textbf{5 of 30}} & \multicolumn{1}{c}{\textbf{}} & \multicolumn{2}{c}{\textbf{Cont. Range Xover}} & \multicolumn{1}{c}{\textbf{13 of 30}} \\ \bottomrule
    \end{tabular}
    \end{table}

@@

    To validate our research, we performed the statistical Z-Test with a ninety-five percent confidence level to analyze the alternatives we present in this proposal. In the ten-dimension category, we found that One Point crossover was the better alternative on four occasions, while our Continuous Range proposal was better on twelve other functions out of thirty. For the thirty-dimensions category, the One Point crossover was better on five occasions, while the Continuous Range crossover was the best alternative on another thirteen. Our experiment results confirm that our solution can be considered a good alternative.

    Some researchers might consider our proposal similar to what is used by algorithms such as Differential Evolution, where we think of our proposal as a variation of existing proposals because we expect this new operator to allow the offspring to continue exploring new search spaces with the birth of individuals. Experimental results indicate that our proposed operator may be a good alternative for the canonical One Point crossover.
